---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
df_with_nas <- read.csv("activity.csv")
# Remove NA rows -- NA only appears in steps column
df <- df_with_nas[complete.cases(df_with_nas),]
df$date <- strftime(df$date, format="%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r message=FALSE}
library(dplyr)
daily <- summarize(group_by(df, date), sum(steps))
names(daily) <- c("date", "steps")
hist(daily$steps,
     col="blue",
     xlab="Daily Steps",
     main="Histogram of Daily Steps",
     breaks=8)

cat("Mean daily steps:", mean(daily$steps))
cat("Median daily steps:", median(daily$steps))
```


## What is the average daily activity pattern?
```{r}
activity_pattern <- df %>%
  group_by(interval) %>%
  summarize(Steps = mean(steps))
plot(activity_pattern, type="l", xlab="Time of day (24 hour clock)")

max_row <- subset(activity_pattern,
                  activity_pattern$Steps==max(activity_pattern$Steps))
cat("The interval with the most steps on average was",
    max_row$interval)
```



## Imputing missing values
```{r}
# Calculate and report the total number of missing values in the dataset
# (i.e. the total number of rows with NAs)
missing_values <- dim(df_with_nas[!complete.cases(df_with_nas),])[1]
cat("Number of rows with missing values:", missing_values)

# Devise a strategy for filling in all of the missing values in the dataset.
# The strategy does not need to be sophisticated. For example, you could use
# the mean/median for that day, or the mean for that 5-minute interval, etc.

# Use the mean for the 5-minute interval
df_imputed <-
  df_with_nas %>% 
  group_by(interval) %>% 
  mutate(steps = ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps))

# Make a histogram of the total number of steps taken each day and Calculate
# and report the mean and median total number of steps taken per day. Do these
# values differ from the estimates from the first part of the assignment? What
# is the impact of imputing missing data on the estimates of the total daily
# number of steps?
daily_imputed <- summarize(group_by(df_imputed, date), sum(steps))
names(daily_imputed) <- c("date", "steps")
hist(daily_imputed$steps,
     col="blue",
     xlab="Daily Steps",
     main="Histogram of Daily Steps",
     breaks=8)

cat("Mean daily steps:", mean(daily_imputed$steps))
cat("Median daily steps:", median(daily_imputed$steps))
```

The mean and median are very close to the previously calculated values. The
histogram shows a higher peak around the mean/median steps per day.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
```

